# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
library(lattice)
data <- read.csv("activity.csv", stringsAsFactors=F)
data$date <- as.Date(data$date)
```

## What is mean total number of steps taken per day?
### A histogram of the total number of steps taken each day
```{r}
sum_data <- aggregate(data$steps~data$date, sum, data=data)
hist(sum_data[,2], breaks=30, main="A histogram of the total number of steps taken each day", xlab="the total number of steps taken each day", col="black")
```
### What is mean total number of steps taken per day?
```{r}
mean(sum_data[, 2])
```
### What is median total number of steps taken per day?
```{r}
median(sum_data[, 2])
```

## What is the average daily activity pattern?
### a time series plot of the 5-minute
```{r}
mean_data <- aggregate(data$steps~data$interval, mean, data=data)
plot(mean_data[,1], mean_data[,2], type="l", xlab="Interval in every 5 minutes", ylab="average daily steps", col="red", main="time series plot of the 5-minute")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
mean_data[mean_data[,2]==max(mean_data[,2]),1]
```

## Imputing missing values
### Calculate and report the total number of missing values in the dataset
```{r}
na_data <-data[is.na(data),]
nrow(na_data)
```
### Fill missing value for that 5-minute interval with mean value, and create a new dataset
```{r}
new_data <- data
for( row in seq(nrow(new_data)) ) {
  if(is.na(new_data[row, 1])) {
    new_data[row, 1] <- mean_data[mean_data[,1] == new_data[row, 3], 2]
  }
}
```
### A histogram of the total number of steps taken each day for new dataset
```{r}
new_sum_data <- aggregate(new_data$steps~new_data$date, sum, data=new_data)
hist(new_sum_data[,2], breaks=30, main="A histogram of the total number of steps taken each day for new dataset", xlab="the total number of steps taken each day", col="black")
```
### What is new dataset's mean total number of steps taken per day?
```{r}
mean(new_sum_data[, 2])
```
### What is new dataset's median total number of steps taken per day?
```{r}
median(new_sum_data[, 2])
```
the values differ from the estimates from the first part of the assignment, because we filled missing value with the average steps for that 5 minutes interval, this change made the data a little different.
## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekday <- weekdays(new_data[,2])
weekday[weekday!="Saturday"&weekday!="Sunday"] <- "weekday"
weekday[weekday=="Saturday"|weekday=="Sunday"] <- "weekend"
weekday <- factor(weekday)
new_data <- cbind(new_data, weekday)
mean_data <- aggregate(steps~interval+weekday, mean, data=new_data)
xyplot(steps ~ interval | weekday, mean_data, type="l",layout=c(1,2), xlab="Interval", ylab="Number of steps")
```